Automated topic naming to support analysis of software maintenance activities

### 2010-08-17 icse/icse2011.tex
Mild edits and todos inserted

\newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
\hyphenation{Max-DB}

@@

\begin{abstract}
  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
 Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
 Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
 However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
 This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
 We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
 Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{abstract}

@@

\section{Introduction}
The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases. It does not take long for a software project to grow to a point where keeping track of these artifacts exceeds easy comprehension. For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself. Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.

Developers do more than merely observe these abstractions, however. They must also use them to assess progress and work with colleagues. To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development artifacts. Such labels can be used to compare progress or understand how well a project is meeting requirements. These labels then become the topics for development conversations (such as the stream of commentary on a bug report). One way to identify topics is to use the artifacts themselves to derive them.

Deriving such topics is difficult. Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. This is impractical when one has to handle more than one hundred different topics. To scale to meaningfully sized projects, we require automatic assistance to determine the topic label.

@@

In this paper, we describe \textbf{Nomen}, which is a tool for \emph{labelled topic extraction}. \textbf{Nomen} addresses two gaps in the topic mining literature:
\begin{enumerate}
  \item Topic mining of software has been limited to one project at a time. This is because traditional topic mining techniques are specific to a particular dataset. \textbf{Nomen} allows for comparisons \textit{between} projects. 
  \item Topic modeling creates bags of words that require interpretation by the user to assign meaning. Like (1), this means
that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels agnostic of project.
\end{enumerate}

@@

\section{Study design and execution}
Figure \ref{fig:process} gives an outline of the methodology followed. We began by gathering source data; then applied unsupervised and supervised learning techniques to identify topics. We then use these topics to analyse the role of non-functional requirements in software maintenance.

@@

\subsection{Generating the data}
\label{sec:wordlist}
To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. We selected systems from the same application domain, so the functional requirements would be broadly similar. We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.  Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software lifecycle.

@@

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

@@

Accordingly, we tried to associate each topic with a label from a list of keywords and related terms. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words (a topic might match several keywords). We used several different word lists for comparison, which we call \textsf{exp1,exp2,exp3} in the text which follows.

@@

@@

\begin{figure*}
  \centering
 \includegraphics[width=.55\textwidth]{unsupervised-bar}
 \caption{ROC values for each NFR, per project and wordlist.}
  \label{fig:maxdb-unsup-results}
\end{figure*}

@@

The less-frequently occurring a label the harder it is to get accurate results, due to the high noise level. Nevertheless, these results are better than our previous word bag results of \textsf{exp2} and \textsf{exp3}, because the ROC values are sufficiently higher in most cases (other than MaxDB reliability and MySQL efficiency). The limitation of the approach we took here is that we assume labels are independent; however, labels could be correlated with each other. We also did not evaluate how well the learners performed together.

\begin{figure*}[ht]
\centering
\subfigure[MySQL]{
\includegraphics[width=0.4\textwidth]{mysql-supervised}
\label{fig:subfig1}
}
\subfigure[MaxDB]{
\includegraphics[width=0.4\textwidth]{maxdb-supervised}
\label{fig:subfig2}
}
\label{fig:best-learn-per-tag}
\caption[]{ROC value (black) for the best learner per label, compared with ZeroR accuracy values (grey).}
\end{figure*}

\subsection{Applying multiple labels to topics}
As noted in Section \ref{sec:wordlist}, each topic in our dataset can be composed of zero to all NFRs. For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement in the source code that related to \textit{usability}. However, traditional machine learning techniques, such as Bayesian networks, can only map topics to a single class. The Mulan~\cite{mulan} library encapsulates several different machine learners which can handle this multi-label learning for our data-set. Mulan also includes methods for determining the performance of such techniques.
% TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank multiple labels at once. We must check if the full subset of labels was applied, and then how much of that subset was applied.

@@

Fig. \ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. %Calibrated Label Ranking (CLR)
The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}. These classifiers performed better than other multi-label classifiers. They have the best micro and macro ROC scores, although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplearn}.
\todo[inline]{discuss this similiarity in more detail}
\todo[inline]{what of CLR? Is it hierarchical?}

\begin{figure*}[ht]
\centering
\subfigure[MySQL]{
\includegraphics[width=0.4\textwidth]{multi-label-results-mysql}
\label{fig:subfig3}
}
\subfigure[MaxDB]{
\includegraphics[width=0.4\textwidth]{multi-label-results-maxdb}
\label{fig:subfig4}
}
\label{fig:mulan}
\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner.}
\end{figure*}

@@

@@

\section{Appendix}
Our data and scripts are available at \url{http://softwareprocess.es/nomen/}

>>>>

\newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
\hyphenation{Max-DB mysql\_-real\_-connect}

@@

\begin{abstract}
\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
  research area and we're saying they don't do enough}
  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
 Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
 Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
 However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
 This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
 We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
 Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{abstract}

@@

\section{Introduction}

The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases. It does not take long for a software project to grow to a point where keeping track of these artifacts exceeds easy comprehension. For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself. Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.

Developers do more than merely observe these abstractions, however. They must also use them to assess progress and work with colleagues. To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development artifacts. Such labels can be used to compare progress or understand how well a project is meeting requirements. These labels then become the topics for development conversations (such as the stream of commentary on a bug report). One way to identify topics is to use the artifacts themselves to derive them. 
\todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up. Topics are used below and it isn't even clear what they are or what we're talking about. Naming is important but lets tell the reader what is being named first.}

Deriving such topics is difficult. Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. This is impractical when one has to handle more than one hundred different topics. To scale to meaningfully sized projects, we require automatic assistance to determine the topic label.
\todo[inline]{Readers still don't know what topics are or what the point is yet..}

@@

In this paper, we describe \textbf{Nomen}, which is a tool for \emph{labelled topic extraction}. \textbf{Nomen} addresses two gaps in the topic mining literature:
\begin{enumerate}
  \item Topic mining of software has been limited to one project at a time. This is because traditional topic mining techniques are specific to a particular data-set. \textbf{Nomen} allows for comparisons \textit{between} projects. 
  \item Topic modeling creates bags of words that require interpretation by the user to assign meaning. Like (1), this means
that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels agnostic of project.
\end{enumerate}

@@

\todo[inline]{Topics are poorly defined by this point and since we're building on research which didn't DO this, we should mention that research. Previous work probably should go here.}

\section{Study design and execution}
Figure \ref{fig:process} gives an outline of the methodology followed. We began by gathering source data; then applied unsupervised and supervised learning techniques to identify topics. We then use these topics to analyze the role of non-functional requirements in software maintenance.

@@

\subsection{Generating the data}
\label{sec:wordlist}
To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. We selected systems from the same application domain, so the functional requirements would be broadly similar. We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.  Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.

@@

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}

@@

Accordingly, we tried to associate each topic with a label from a list of keywords and related terms. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words (a topic might match several keywords). We used several different word lists for comparison, which we call \textsf{exp1, exp2, exp3} in the text which follows.

@@

\todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

\begin{figure*}
  \centering
 \includegraphics[width=.55\textwidth]{unsupervised-bar}
 \caption{ROC values for each NFR, per project and wordlist.}
\todo[inline]{this is too tiny in print}

  \label{fig:maxdb-unsup-results}
\end{figure*}

@@

The less-frequently occurring a label the harder it is to get accurate results, due to the high noise level. Nevertheless, these results are better than our previous word bag results of \textsf{exp2} and \textsf{exp3}, because the ROC values are sufficiently higher in most cases (other than MaxDB reliability and MySQL efficiency). The limitation of the approach we took here is that we assume labels are independent; however, labels could be correlated with each other. We also did not evaluate how well the learners performed together.
\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}

\begin{figure*}[ht]
\centering
\subfigure[MySQL]{
\includegraphics[width=0.4\textwidth]{mysql-supervised}
\label{fig:subfig1}
}
\subfigure[MaxDB]{
\includegraphics[width=0.4\textwidth]{maxdb-supervised}
\label{fig:subfig2}
}
\label{fig:best-learn-per-tag}
\caption[]{ROC value (black) for the best learner per label, compared with ZeroR accuracy values (grey).
\todo[inline]{better discussion and graphical legend would be nice}
}
\end{figure*}

\subsection{Applying multiple labels to topics}
As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero to all NFRs. For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement in the source code that related to \textit{usability}. However, traditional machine learning techniques, such as Bayesian networks, can only map topics to a single class. The Mulan~\cite{mulan} library encapsulates several different machine learners which can handle this multi-label learning for our data-set. Mulan also includes methods for determining the performance of such techniques.
% TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank multiple labels at once. We must check if the full subset of labels was applied, and then how much of that subset was applied.

@@

Fig. \ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. %Calibrated Label Ranking (CLR)
The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}. These classifiers performed better than other multi-label classifiers. They have the best micro and macro ROC scores, although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplearn}.
\todo[inline]{discuss this similarity in more detail}
\todo[inline]{what of CLR? Is it hierarchical?}

\begin{figure*}[ht]
\centering
\subfigure[MySQL]{
\includegraphics[width=0.4\textwidth]{multi-label-results-mysql}
\label{fig:subfig3}
}
\subfigure[MaxDB]{
\includegraphics[width=0.4\textwidth]{multi-label-results-maxdb}
\label{fig:subfig4}
}
\label{fig:mulan}
\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner.
\todo[inline]{better discussion and graphical legend would be nice}
}
\end{figure*}

@@

\todo[inline]{Update this}

@@

%XXX can save space here!
\section{Appendix}
Our data and scripts are available at \url{http://softwareprocess.es/nomen/}
